Nondeterministic Bigraphical Reactive Systems for Markov Decision Processes\thanks{Supported by organization x.}

### 2018-08-06 paper.tex
Describing the first agent (almost done)

@@

@@

From \cite{dblp:journals/tcs/sevegnanic15}.

@@

\begin{definition}[concrete bigraph with sharing]
  A concrete bigraph
  \[ F = (V_F, E_F, \ctrl_F, \prnt_F, \link_F) : \langle k, X \rangle \to
    \langle m, Y \rangle \]
  consists of a concrete place graph with sharing $F^P = (V_F, \ctrl_F, \prnt_F)
  : k \to m$ and a concrete link graph $F^L = (V_F, E_F, \ctrl_F, \link_F) : X
  \to Y$. If $X = \epsilon$, then $F$ is called \emph{ground}. We write the
  concrete bigraph with sharing as $F = (F^P, F^L)$.
\end{definition}

From PhD thesis \cite{DBLP:phd/ethos/Sevegnani12}

\begin{definition}[reaction rule] \label{reaction_rule}
  A \emph{reaction rule} is a pair
  \[ \mathsf{R} = (R : m \to J, R' : m \to J), \]
  sometimes written as $R \longrightarrowRHD R'$, where $R$ is the \emph{redex}
  and $R'$ the \emph{reactum}, and $R$ is solid. The rule generates all the
  \emph{ground reaction rules} $(r, r')$, where $r = (R \otimes \id_Y) \circ d$
  and $r' = (R' \otimes \id_Y) \circ d$ for some discrete ground
  \emph{parameter} $d : \epsilon \to \langle m, Y \rangle$. The \emph{reaction
    relation} $\longrightarrowrhd_{\mathsf{R}}$ over ground bigraphs is defined
  by
  \[ g \longrightarrowrhd_{\mathsf{R}} g' \text{ iff } g = Dr \text{ and } g' = Dr' \]
  for some bigraph $D$ and some ground reaction rule $(r, r')$ generated from
  $\mathsf{R}$.
\end{definition}

\begin{definition}[bigraphical reactive system (BRS)]
  A \emph{bigraphical reactive system} consists of a pair $(\mathcal{B},
  \mathcal{R})$, where $\mathcal{B}$ is a set of bigraphs and $\mathcal{R}$ is a
  set of reaction rules defined over $\mathcal{B}$. It has a reaction relation
  \[ \longrightarrowrhd_{\mathcal{R}} \coloneqq \bigcup_{\mathsf{R} \in
      \mathcal{R}} \longrightarrowrhd_{\mathsf{R}}, \]
  which will be written $\longrightarrowrhd$ when $\mathcal{R}$ is understood.
\end{definition}

\subsection{Probabilistic bigraphs}

@@

\begin{definition}[probabilistic bigraphical reactive system (PBRS)]
  A \emph{probabilistic bigraphical reactive system} consists of a pair
  $(\mathcal{B}, \mathcal{R})$, where $\mathcal{B}$ is a set of bigraphs and
  $\mathcal{R}$ is a set of probabilistic reaction rules defined over
  $\mathcal{B}$.

@@

\subsection{Nondeterministic bigraphs}

@@

\begin{definition}[nondeterministic bigraphical reactive system (NBRS)]
  A \emph{nondeterministic bigraphical reactive system} consists of a pair
  $(\mathcal{B}, \mathcal{R})$, where $\mathcal{B}$ is a set of bigraphs and
  $\mathcal{R}$ is a set of nondeterministic reaction rules defined over
  $\mathcal{B}$.

@@

% TODO: states
BigraphER \cite{DBLP:conf/cav/SevegnaniC16} supports defining predicates, i.e.,
bigraphs that are compared to every encountered state. We can associate a reward
with each predicate, allowing us to assign rewards to states in a flexible and
semantically meaningful way: the reward of a state is simply the sum of the
rewards of all matching predicates (and $0$ in case there are none).

@@

\begin{figure}
  \includegraphics{images/interface.jpg}
  \caption{The BigraphER Jupyter interface with syntax highlighting.}
  \label{interface}
\end{figure}

\begin{figure}
  \includegraphics{images/transitions.jpg}
  \caption{...}
  \label{transitions}
\end{figure}

\begin{figure}
  \includegraphics{images/simulation.jpg}
  \caption{...}
  \label{simulation}
\end{figure}

\begin{figure}
  \includegraphics{images/output.jpg}
  \caption{...}
  \label{output}
\end{figure}

\section{Case study in autonomous agents}

>>>>

\ttfamily
\DeclareFontShape{OT1}{lmtt}{m}{it}
     {<->sub*lmtt/m/sl}{}

@@

\subsection{Markov Decision Processes}

@@

\subsection{Bigraphs}

From \cite{dblp:journals/tcs/sevegnanic15}.
From PhD thesis \cite{DBLP:phd/ethos/Sevegnani12}

\begin{figure}
  \includegraphics{models/example_bigraph/foo_lhs.pdf}
  \caption{An example bigraph.}
  \label{example_bigraph}
\end{figure}

Consider the graphical representation of a bigraph in Fig.
\ref{example_bigraph}. The two white dashed rectangles represent \emph{regions},
i.e. parents of any otherwise parentless nodes. Grey dashed rectangles are
called \emph{sites} and encode parts of the model that have been abstracted
away. Nodes, sites, and regions are also known as \emph{places}. Two places with
the same parent, or two points with the same link are called \emph{siblings}.

@@

\begin{definition}[concrete bigraph with sharing]
  A concrete bigraph
  \[ F = (V_F, E_F, \ctrl_F, \prnt_F, \link_F) : \langle k, X \rangle \to
    \langle m, Y \rangle \]
  consists of a concrete place graph with sharing $F^P = (V_F, \ctrl_F, \prnt_F)
  : k \to m$ and a concrete link graph $F^L = (V_F, E_F, \ctrl_F, \link_F) : X
  \to Y$. We write the concrete bigraph with sharing as $F = (F^P, F^L)$.
\end{definition}

Given a set of reaction rules, we refer to the configurations that a system may
adopt as \emph{states}. A bigraph with inner face $\epsilon$ is called
\emph{ground}.

\begin{definition}[solid bigraph]
  A bigraph is \emph{solid} if these conditions hold:
  \begin{enumerate}
  \item no regions or outer names are idle (a region is called \emph{idle}
    if it is empty);
  \item no two sites or inner names are siblings;
  \item the parent of every site is a node;
  \item no outer name is linked to an inner name.
  \end{enumerate}
\end{definition}

\begin{definition}[reaction rule] \label{reaction_rule}
  A \emph{reaction rule} is a pair
  \[ \mathsf{R} = (R : m \to J, R' : m \to J), \]
  sometimes written as $R \longrightarrowRHD R'$, where $R$ is the \emph{redex}
  and $R'$ the \emph{reactum}, and $R$ is solid. The rule generates all the
  \emph{ground reaction rules} $(r, r')$, where $r = (R \otimes \id_Y) \circ d$
  and $r' = (R' \otimes \id_Y) \circ d$ for some discrete ground
  \emph{parameter} $d : \epsilon \to \langle m, Y \rangle$. The \emph{reaction
    relation} $\longrightarrowrhd_{\mathsf{R}}$ over ground bigraphs is defined
  by
  \[ g \longrightarrowrhd_{\mathsf{R}} g' \text{ iff } g = D \circ r \text{ and
    } g' = D \circ r' \]
  for some bigraph $D$ and some ground reaction rule $(r, r')$ generated from
  $\mathsf{R}$.
\end{definition}

\begin{definition}[bigraphical reactive system (BRS)]
  A \emph{bigraphical reactive system} consists of a pair $(\mathcal{B},
  \mathcal{R})$, where $\mathcal{B}$ is a set of ground bigraphs and
  $\mathcal{R}$ is a set of reaction rules defined over $\mathcal{B}$. It has a
  reaction relation
  \[ \longrightarrowrhd_{\mathcal{R}} \coloneqq \bigcup_{\mathsf{R} \in
      \mathcal{R}} \longrightarrowrhd_{\mathsf{R}}, \]
  which will be written $\longrightarrowrhd$ when $\mathcal{R}$ is understood.
\end{definition}

\section{Extensions to Bigraphs}

\subsection{Probabilistic Bigraphs}

@@

\begin{definition}[probabilistic bigraphical reactive system (PBRS)]
  A \emph{probabilistic bigraphical reactive system} consists of a pair
  $(\mathcal{B}, \mathcal{R})$, where $\mathcal{B}$ is a set of ground bigraphs
  and $\mathcal{R}$ is a set of probabilistic reaction rules defined over
  $\mathcal{B}$.

@@

\subsection{Nondeterministic Bigraphs}

@@

\begin{definition}[nondeterministic bigraphical reactive system (NBRS)]
  A \emph{nondeterministic bigraphical reactive system} consists of a pair
  $(\mathcal{B}, \mathcal{R})$, where $\mathcal{B}$ is a set of ground bigraphs
  and $\mathcal{R}$ is a set of nondeterministic reaction rules defined over
  $\mathcal{B}$.

@@

BigraphER \cite{DBLP:conf/cav/SevegnaniC16} supports defining predicates, i.e.,
bigraphs that are compared to every encountered state. We can associate a reward
with each predicate, allowing us to assign rewards to states in a flexible and
semantically meaningful way: the reward of a state is simply the sum of the
rewards of all matching predicates (and $0$ in case there are none).

@@

% TODO: introduce this
\begin{figure}
  \includegraphics{images/interface.jpg}
  \caption{The BigraphER Jupyter interface with syntax highlighting.}
  \label{interface}
\end{figure}

\section{A Case Study in Autonomous Agents}

Akin to \cite{dblp:conf/nfm/giaquintahimn18}.

\subsection{\#1}

Grid layout (but could be any nodes-and-edges layout, as long as it's clear how
to go back home). Tracking visited/unvisited states. Each unvisited state has a
probability $p$ of containing an object. The agent returns (and stays) home
after collecting a set number of objects (easily customisable). Starting at the
bottom left cell.

\lstinputlisting[caption=Controls., captionpos=b, language=Big,
label=agent1_controls, firstline=3, lastline=17]{models/agent1.big}

We begin by introducing controls: the types of nodes, what they contain, and
what they are linked to. The full list of controls is in Listing
\ref{agent1_controls}. \texttt{Cell} represents a discrete location. It is
linked to either \texttt{Visited} or \texttt{Unvisited} (initially, all except
one cells are unvisited). \texttt{Cell}s always contain
\texttt{Directions}, and one \texttt{Cell} contains an \texttt{Agent}. The
\texttt{Agent} may contain an \texttt{Object}, which may contain another
\texttt{Object}, etc. \texttt{Directions} always contain four nodes:
\texttt{North}, \texttt{East}, \texttt{West}, \texttt{South}. A \texttt{West}
node can be linked either to one of the two \texttt{Void}s or to an
\texttt{East} node of a neighbouring \texttt{Cell} (and similarly with other
directions). \texttt{Void}s represent the edges of the map. Using two
\texttt{Void} controls prevents us from unintentionally connecting the opposite
edges of the map. \texttt{Void1} corresponds to top and left edges, while
\texttt{Void2} is for bottom and right.

\lstinputlisting[caption=Predicates and the initial state., captionpos=b,
language=Big, label=agent1_code, firstline=19,
lastline=50]{models/agent1.big}

Next, we define two predicates: \texttt{home} and \texttt{goal} (see Listing
\ref{agent1_code}). The former states that the \texttt{Agent} is home
if it is in the southwest corner of the grid,
while the latter says that the agent should go home as soon as
it gets one object. A larger number of objects can be set by,
e.g., \texttt{Agent.Object.Object}. Additionally, Listing \ref{agent1_code}
defines the initial $2 \times 2$ grid, as described previously.

\lstinputlisting[caption=The action of going \texttt{north}, captionpos=b,
language=Big, label=agent1_north, firstline=108,
lastline=141]{models/agent1.big}

Lastly, we have five actions (four for the directions, and one for
going home). Each of the actions for directions carries three reaction
rules: one for going to an already visited cell, one for visiting a
cell for the first time, and one for visiting a cell for the
first time and finding an object. We will use the rules for going north
in Listing \ref{agent1_north} as an example. When going to an already visited
cell, the \texttt{Agent} simply traverses a link between the \texttt{North} node
of its previous \texttt{Cell} and the corresponding \texttt{South} node. As this
is the only reaction rule for this situation, its probability is $1$. When
going to an unvisited node, we equip an \texttt{Object} with probability $p$.
Regardless of whether that happens or not, the \texttt{Cell} changes its
status from \texttt{Unvisited} to \texttt{Visited}. % TODO: going home now

\section{Conclusion}
